### 2015-03-04 paper/conclusion.tex
updating the paper

\section{Conclusions and Future Work}
In this paper we have described our approach to achieving high-throughput multi-stage stream processing.  We have used application level message buffering to improve serialization and deserialization process and thread pools to improve parallelism.

@@

Our future work will mainly target two areas: reliable message delivery and compression. We will focus on implementing a reliable message delivery protocol to support at least one grantee on top of our communication framework.  The work on compression will target reducing the overthe-wire footprints of the packets. We will explore the use of compression algorithms while ensuring that the gains in message size reduction do not result in unacceptable latency overheads and also that the throughput improves.

>>>>

\section{Conclusions and Future Work}
In this paper we have described our approach to achieving high-throughput multi-stage stream processing.  We have used application message buffering to improve serialization and deserialization process and thread pools to improve parallelism.

@@

Our future work will mainly target two areas: reliable message delivery and compression. We will focus on implementing a reliable message delivery protocol to support at least one guarantee on top of our communication framework.  The work on compression will target reducing the overthe-wire footprints of the packets. We will explore the use of compression algorithms while ensuring that the gains in message size reduction do not result in unacceptable latency overheads and also that the throughput improves.

### 2015-03-04 paper/experiments.tex
updating the paper

\subsection{Effect of message buffer size and thread pool}
For this experiment, we used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data. The EventProducer reads a file containing over 7500000
 ECG records and pushes events with multiple threads. EventReceiver receives the ECG events, processes them and calculates heart rate interval periodically using Pan Tompkins algorithm.

@@

We used the same graph (Figure \ref{ecgGraph}) used in above experiment with 1,2 and 4 \textit{EventReciver} nodes with one \textit{EventSender} node for this experiment. Since 40 thread pool size and 40 message buffer size gives highest throughput, we used that values with our solution. Further we compared our system with Yahoo S4 \cite{neumeyer2010s4} and Twitter Storm \cite{toshniwal2014storm} with the same graph. Twitter Storm \cite{toshniwal2014storm} places tasks arbitrarily within the available nodes. Therefore we used 1 task for \textit{EventSender} and 1,2 and 4 tasks for \textit{EventReceivers} to place them in different nodes. This makes Twitter Storm \cite{toshniwal2014storm} execute only one thread to process data in each node. In order to compare this situation, we measured our system performance with 1 thread and 10 message buffer size as well.

@@

\subsection{Multistage Graph Processing}
We performed a multistage graph test for our system using the 3-level graph shown in Figure \ref{multigraph}. The data as well as the processing logic were obtained from the Grand Challenge problem at 8th ACM International Conference on Distributed Event Based Systems \cite{acm}. This project explores the possibility of using stream processing to process data generated in the smart grid domain. The data is generated by sensors attached to a device called “smart plug” which is plugged as an intermediate layer between a power outlet and the plug of a home appliance. The sensors capture various measures related to power consumption of each connected device and send these measurements continuously to a central processing system. Each household contains several of such smart plugs which will report the sensor readings once in every second. At the central processing system, these data should be processed in order to forecast the short time load and to generate load statistics for real time demand management. We implemented the first part of this project which predicts the load using a machine learning algorithm. This algorithm divides the events to fixed intervals called  time slices and uses average value of time slices to predict the average of feature time slices. We used the publicly available 500MB of data to generate the events.

@@

\subsection{Scalability and Latency}
We analyze the scalability of our framework using cumulative throughput with a similar experiments found in research literature.  In this experiment a given set of nodes communicate directly with each other using TCP connections.  We used same message type (46 bytes in serialized format) used for earlier experiment. The given cumulative throughput is at the application level message rate omitting the TCP overhead and all communications are through the network. For all experiments we use 20 message buffer.

@@

As shown in Figure \ref{allthroughput} system linearly scales with the number of nodes.  Compared to Naiad (Figure 6(a) in cited paper) \cite{murray2013naiad} and Spark Streaming (Figure 4 in cited paper) \cite{zaharia2013discretized} we have achieved better performance.  Figure \ref{alllatency} shows the mean message latency for all messages. We measured the message latency during the same experiment by sending the time stamp along with each message. Compared to Spark Streaming (Figure 4 in cited paper) \cite{zaharia2013discretized} we have achieved much lesser latency level. Naiad (Figure 6(b) in cited paper) \cite{murray2013naiad} reports very low latency for Global barrier messages. However our latency corresponds to message latency for data messages. MillWheel (Section 8.1) \cite{akidau2013millwheel} has reported very low latency level compared to our results. Further it is worth note that the latency can be reduced by reducing the buffer size and hence reducing the throughput. We choose 20 message buffer since it gives better throughput as well as comparable low latency.

>>>>

\subsection{Effect of message buffer size and thread pool}
For this experiment, we used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data. The EventProducer reads a file containing over 7500000
 ECG records and pushes events with multiple threads. EventReceiver receives the ECG events, processes them and calculates heart rate interval periodically using Pan Tompkins algorithm \cite{kohler2002principles}.

@@

We used the same graph (Figure \ref{ecgGraph}) used in above experiment with 1,2 and 4 \textit{EventReciver} nodes with one \textit{EventSender} node for this experiment. Since 40 thread pool size and 40 message buffer size gives highest throughput, we used that values with our solution (named as Granules). Further we compared our system with Yahoo S4 \cite{neumeyer2010s4} and Twitter Storm \cite{toshniwal2014storm} with the same graph. Twitter Storm \cite{toshniwal2014storm} places tasks arbitrarily within the available nodes. Therefore we used 1 task for \textit{EventSender} and 1,2 and 4 tasks for \textit{EventReceivers} to place them in different nodes. This makes Twitter Storm \cite{toshniwal2014storm} execute only one thread to process data in each node. In order to compare this situation, we measured our system performance with 1 thread and 10 message buffer size as well.

@@

\subsection{Multistage Graph Processing}
We performed a multi-stage graph test for our system using the 3-level graph shown in Figure \ref{multigraph}. The data as well as the processing logic were obtained from the Grand Challenge problem at 8th ACM International Conference on Distributed Event Based Systems \cite{acm}. This project explores the possibility of using stream processing to process data generated in the smart grid domain. The data is generated by sensors attached to a device called “smart plug” which is plugged as an intermediate layer between a power outlet and the plug of a home appliance. The sensors capture various measures related to power consumption of each connected device and send these measurements continuously to a central processing system. Each household contains several of such smart plugs which will report the sensor readings once in every second. At the central processing system, these data should be processed in order to forecast the short time load and to generate load statistics for real time demand management. We implemented the first part of this project which predicts the load using a machine learning algorithm. This algorithm divides the events to fixed intervals called  time slices and uses average value of time slices to predict the average of feature time slices. We used the publicly available 500MB of data set to generate the events.

@@

\subsection{Scalability and Latency}
We analyze the scalability of our framework using cumulative throughput with a similar experiments found in research literature \cite{murray2013naiad} \cite{zaharia2013discretized}.  In this experiment a given set of nodes communicate directly with each other using TCP connections.  We used same message type (46 bytes in serialized format) used for earlier experiment. The given cumulative throughput is at the application level message rate omitting the TCP overhead and all communications are through the network. For all experiments we use 20 message buffer. Since each node is capable of sending and receiving 1Gbps of data, ideal throughput is same as number of nodes times 1Gbps.

@@

As shown in Figure \ref{allthroughput} system linearly scales with the number of nodes.  Compared to Naiad (Figure 6(a) in cited paper) \cite{murray2013naiad} and Spark Streaming (Figure 4 in cited paper) \cite{zaharia2013discretized} we have achieved better performance (comparing the values given in those papers).  Figure \ref{alllatency} shows the mean message latency for all messages. We measured the message latency during the same experiment by sending the time stamp along with each message. Compared to Spark Streaming (Figure 4 in cited paper) \cite{zaharia2013discretized} we have achieved much lesser latency level. Naiad (Figure 6(b) in cited paper) \cite{murray2013naiad} reports very low latency for global barrier messages. However our latency corresponds to message latency for data messages. MillWheel (Section 8.1) \cite{akidau2013millwheel} has reported very low latency level compared to our results. Further it is worth note that the latency can be reduced by reducing the buffer size and hence reducing the throughput. We choose 20 message buffer since it gives better throughput as well as comparable low latency.

### 2015-03-04 paper/introduction.tex
updating the paper

Multi-stage, distributed stream processing is useful in analyzing data streams generated by programs, sensors, or users in real time. Analysis of click-streams, tweets, and stock-quotes are primary examples where such processing is often performed. Most commonly, in such systems packets encapsulate tuples representing values corresponding to a set of variables.

These systems receive billions of tuples daily expecting results within fractions of second. Hence these systems are required to process data with high throughput and low latency. The processing logic of these systems are expressed as multi stage graphs and deployed into clusters of hundreds of commodity machines. Each stage receive messages from the previous stage, process them according to a user defined logic and forward them to next stage. Therefore performance of the inter node communication make a huge impact of overall system performance.

Apart from these primary requirements, other features such as reliable message delivery,  state preservation and load balancing are also required depending on the nature of the high level applications. There have been many frameworks \cite{toshniwal2014storm} \cite{zaharia2010spark} \cite{akidau2013millwheel} \cite{murray2013naiad} developed to address these issues. Further all these research work focus on providing better computational models to develop these systems encapsulating the communication complexity from the user as well.  However these researches have paid less attention to inter node communication efficiency. Therefore in this paper we focus on this inter node communication aspect and identify means of increasing throughput while keeping low latency levels.

First we started research on parallel TCP connections \cite{hacker2001end} to improve the performance since it has been used in wide area network applications. Although parallel TCP improves throughput of the system, we found that even without that it is possible to saturate the network link (1 Gbps link connected through a switch) with 256 byte array messages. However user APIs of the modern stream processing systems, deals with programming language objects rather than byte array messages. Therefore we can improve the performance of node communication by increasing the language level object to byte array conversion process. Further serialization form of these language objects (if proper serialization techniques used) can be small as well. Hence buffering programing level objects and serialize them as batches would greatly improve the performance.

As a contribution of this paper we introduce those concepts and provide a reference implementation to compare with other frameworks and analyses the scalability and latency. These general concepts can be applied to any other software to improve its performance.

The remainder of this paper is organized as follows: Section 2 provides an overview of our system at user level. Section 3 illustrates the underlying design of the system. Section 4 analyses the experiment results which demonstrate the performance of our solution. Section 5 discusses the related work and finally section 6 provides conclusions and future work.

>>>>

Multi-stage, distributed stream processing is useful in analyzing data streams generated by programs, sensors, or users. Analysis of click-streams, tweets, and stock-quotes are primary examples where such processing is often performed. Most commonly, in such systems packets encapsulate \textit{tuples} representing values corresponding to a set of variables. One advantage of doing multi-stage processing is that individual stages can be scaled horizontally in response to the load i.e., each stage could have multiple instances.

The stages involved in such stream processing can be fluid.  It is possible for stages to be added and removed dynamically. Furthermore, different processing pipelines may share stages. Along the same lines, a data packet may be processed within multiple processing pipelines. Individual stages may transform the packets being processed before forwarding it to other stages. Stream processing systems do not place restrictions on the type of packet transformations and modifications that can be performed.

The rates at which the stream packets arrive place unique strains on such systems. Each packet results in a mix of processing and I/O at each stage. Failure to keep up with the data generation rates result in queue build-ups, followed by overflows, and subsequent process failures. Furthermore, in the case of a multi-stage processing pipeline, the processing is only as fast as the slowest stage in the system.

There are two key aspects in processing stream packets: latency and throughput. Latency corresponds to the end-to-end delay as the packet makes its way through the processing pipeline. This metric is useful in characterizing the how timely the processing is. Throughput is a measure of how many packets can be processed per-second within a pipeline. The throughput represents how well the system can cope with rates of data arrivals. As the number of processing stages increase we expect to achieve higher throughput, though latency may increase a little corresponding to the number of hops within the pipeline.

\subsection{Research Challenges} 
In this paper we consider the problem of designing a scalable framework for the high throughput processing of data streams. There are several challenges in achieving this.
\begin{enumerate}
	\item Continuous data arrivals: In the systems we consider data is continually arriving at high rates. Inability to keep pace with the arrival rates will result in buffer overflows.
	\item Shared communication links: Stages comprising the stream processing pipeline may be distributed over multiple machines, and the links connecting these stages are shared Ethernet LANs. Effective utilization of these links is important for achieving high throughput.
	\item Commodity machines: Individual stages execute on commodity machines that have limited memory (order of a few GB) and processing cores (4).  So there are limits to the gains that can be accrued by processing a single packet faster.
\end{enumerate}

\subsection{Research Questions}
Support for high-throughput stream processing involves accounting for aspects relating to memory and processing at individual stages and also for communications between stages comprising the pipeline. Specific research questions that we explore include:
\begin{enumerate}
	\item How can we effectively manage memory consumption during processing? During packet processing, tuples must be extracted and processed. This involves allocation and garbage collection of memory. Given the rate at which packets arrive, operations relating memory management must be managed effectively.
	\item How can we ensure effective management of the processing workloads? Since each packet is processed independently, the framework must allow for multiple packets to be processed concurrently.
	\item How can we reduce latencies involved during processing? Since message passing between stages involves network I/O, we need to ensure that these I/O operations can be effectively interleaved with processing. Inefficiencies in interleaving result in serial processing that may cause queue build-ups.
\end{enumerate}

\subsection{Approach Summary}

The work described here provides a framework for multistage stream processing. Users are only required to specify the processing graph, connectivity between individual stages, and the processing to be performed by each stage. Our framework manages all aspects relating to memory management, efficiencies in communications, and concurrent processing.

Effective memory management is important to avoid buffer overflows and out-of-memory errors. We achieve this by reusing objects. Our approach precludes the need to create an object for every packet arrival. Given the rates at which data packets arrive, memory management costs relating to managing the object lifecycle (creation, initialization, and garbage collection) would be prohibitive. An additional artifact in creation of objects per packet is thrashing due to page faults as the memory consumption at a stage increases.

Given that the stages are dispersed, message passing between stages comprising the pipeline must not be inefficient. Our effort targets effective utilization of the underlying network in two ways. First, we buffer application messages (messages produced by application logic). This buffering targets minimizing the creation of runtime objects for a large number messages while generating large enough messages to utilize the underlying network. The amount of messages get buffered can be configured to balance the competing pulls of timely delivery and high-throughput transfer. Second, our serialization scheme allows for compact over-the-wire representation of the tuples within the packets. Inefficiencies in serializations schemes result in verbose representations that increase the network footprint of individual packets.

Our framework also focuses on efficiencies in processing these packets. There are two ways in which we accomplish this. First, we rely on thread pools at each stage. Packet processing is performed within threads and the thread-pool is sized so that we balance concurrency gains and thread-switching overheads. Second, we harness non-blocking I/O. Given that packet processing at each stage involves network I/O the use of non-blocking I/O allows us to interleave processing and I/O much more effectively.

We evaluate the suitability of our methodology by profiling its performance with real applications. This includes real-time ECG processing where we are analyzing waveform data in real-time to perform QRS-complex detection using the well-known Pan-Tompkins \cite{kohler2002principles} algorithm and a multi-stage query to analyze smart plug data. Finally, we also contrast the performance of our system with systems such as Twitter Storm \cite{toshniwal2014storm} and  Yahoo S4 \cite{neumeyer2010s4} that do not incorporate some of the efficiencies in our system. Our results validate the suitability of our approach with substantially higher throughputs that what can be achieved in these systems. Furthermore, several aspects of our methodology can be incorporated into these aforementioned systems to improve their performance as well.

\subsection{Paper Contributions}
In this paper we have described how to achieve high-throughput, multi-stage distributed stream processing. Our specific contributions include the following:
\begin{enumerate}
	\item Our approach identifies aspects that play a key role in achieving high-throughput processing. High-throughput processing is possible by taking a holistic view of the system encompassing memory, computing, and communications efficiencies. This paper demonstrates how combining object reuse, serialization efficiencies, concurrent processing, buffering application messages, and the use of non-blocking I/O allows us to achieve high throughput. 
	\item The per-packet processing latencies that we achieve in our system demonstrate suitability for applications where such timely processing is necessary.
	\item We provide a simple framework for user to express their processing graph. Users only need to specify this graph and the processing that must be performed for each stage. Users are freed from the dealing with concurrent, I/O, and memory efficiency issues. 
	\item Our results demonstrate the suitability of our approach. We have contrasted the performance of our system with Twitter Storm \cite{toshniwal2014storm} and  Yahoo S4 \cite{neumeyer2010s4}.
\end{enumerate}

\subsection{Paper Organization}
The remainder of this paper is organized as follows: Section 2 provides an overview of our system at user level. Section 3 illustrates the underlying design of the system. Section 4 analyses the experiment results which demonstrate the performance of our solution. Section 5 discusses the related work and finally section 6 provides conclusions and future work.

### 2015-03-04 paper/system_architecture.tex
modifing introduction and conclusions
updating the paper

\section{System Design}
In this section we describe key aspects of our system design which increase the throughput of inter node communication. First of all it is worth to note that even with one TCP connection, it is possible to saturate the network link usage, if the sending and receiving processes send and receive byte arrays without any delay even with 256 bytes array messages. However serializing java object messages to byte arrays and deserializing byte arrays back to java objects take considerable amount of time. Hence the key aspect of  improving the system performance is to reduce this message processing delay. We use multitasking and bulk message serializing to solve this problem.

In this design, we use thread pools both at client side and server side to processes messages while sharing a single TCP connection. Firstly both thread pools access underline connection resource only to send receive byte arrays making message serialization and deserialization process parallel. Secondly it serializes messages together to improve performance of the serialization process.  Figure \ref{interprocess} shows the design involved in sending a message from one process to another.

Lets assume client side process requires to send a set of messages to server side process. In order to send these messages first the client process threads send these messages to its \textit{ElementContainer}. \textit{ElementContainer} holds all the streams (this is an abstraction of a link in the process graph) to which these messages need to be send and it pushes each message to all the streams. Stream decides the target node and buffers messages for a configured amount of size. Then it passes the buffered  message list along the target node details to \textit{ConnectionManager} which holds \textit{ClientConnection}s for each node. \textit{ConnectionManager} serializes the message list and sends the byte array to correct \textit{ClientConnection} using the target node. \textit{ClientConnection} uses the \textit{DataOutput} which wraps the TCP connection to send message to server side.

At the server side there is a set of \textit{ServerTask}s which reads receiving messages from a pool of \textit{DataInput} streams available in the \textit{ServerConnection} (we register these connections at the connections creating time) and send them to receiving processes. First a server task acquires a \textit{DataInput} stream and reads the binary message and release it. Then it deserializes the binary message using the sending process id to identify the event type.  After that it passes this message to the \textit{WorkerContainer} which holds all processes. Finally \textit{WorkerContainer} dispatches this message to correct process using receiving process id. Message ordering happens at the process level if required.

\begin{figure}[!tii]
        \centering
        \includegraphics[width=3.0in]{interprocess.png}
        \caption{Communication between two Process}
        \label{interprocess}
\end{figure}

>>>>

\section{System Design}
In this section we describe key aspects of our system design which increase the throughput of inter node communication and underline framework implementation details. First of all it is worth to note that even with one TCP connection, it is possible to saturate the network link usage, if the sending and receiving processes send and receive byte arrays without any delay even with 256 bytes array messages. However serializing java object messages to byte arrays and deserializing byte arrays back to java objects take considerable amount of time. Hence the key aspect of  improving the system performance is to reduce this message conversion delay. We use multitasking and bulk message serializing to solve this problem.

In this design, we use thread pools both at client side and server side to process messages while sharing a single TCP connection. Firstly both thread pools access underline connection resource only to send and receive byte arrays making message serialization and deserialization process parallel. Secondly it buffers the messages and serializes together.  This makes sure it produces large enough binary messages to fully utilize the underlying network and minimize the number of runtime objects created for message conversion. Figure \ref{interprocess} shows the design involved in sending a message from one process to another.

Let's assume client side process requires to send a set of messages to server side process. In order to send these messages, first the client process threads send these messages to its \textit{ElementContainer}. \textit{ElementContainer} holds all the streams (this is an abstraction of a link in the process graph) to which these messages need to be send and it pushes each message to all the streams. Stream decides the target node and buffers messages for a configured amount of size. Then it passes the buffered  message list along the target node details to \textit{ConnectionManager} which holds \textit{ClientConnection}s for each node. \textit{ConnectionManager} serializes the message list and sends the byte array to correct \textit{ClientConnection} using the target node. \textit{ClientConnection} uses the \textit{DataOutput} which wraps the TCP connection to send message to server side.

At the server side there is a set of \textit{ServerTask}s which reads receiving messages from a pool of \textit{DataInput} streams available in the \textit{ServerConnection} (we register these connections at the connections creating time) and send them to appropriate processes. First a server task acquires a \textit{DataInput} stream and reads the binary message and release it. Then it deserializes the binary message using the sending process id to identify the event type.  After that it passes this message to the \textit{WorkerContainer} which holds all processes. Finally \textit{WorkerContainer} dispatches this message to correct process using receiving process id. Message ordering happens at the process level if required.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{interprocess.png}
        \caption{Communication between two Process}
        \label{interprocess}
\end{figure}

\subsection{Application message processing}
We use \textit{DataOutputStream} with \textit{ByteArrayOutputStream} to serialize application messages to binary format as batches and \textit{DataInputStream} with \textit{ByteArrayInputStream} to deserialize from the binary format. First four bytes of the binary format is used to specify the number of messages in the batch.  Deserializer use this information to read the exact amount of messages from the binary message. As mentioned earlier, we minimize these object creation by processing messages together. We also experimented with Kryo \cite{kryo}. Kryo provides an efficient binary format in terms of number of bytes used but that causes less throughput compared to above mechanism.

\subsection{Binary Message Communication}
Once the application messages are converted to binary format, binary messages need to be send to the server side. At the client side we use \textit{DataOutput} to write the message length with the message and at the server side, we use \textit{DataInput} to read messages according to the length. \textit{DataOutput} and \textit{DataInpu}t interfaces requires \textit{OutputStream} and \textit{InputStream} objects to write and read data. We implement those interfaces with java non blocking I/O API as given bellow. This implementation allow as to re use the underline \textit{ByteBuffer} objects.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{client.png}
        \caption{OutputStream interface implementation with java blocking API}
        \label{client}
\end{figure}
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{server.png}
        \caption{InputStream interface implementation with java blocking API}
        \label{server}
\end{figure}

At client side (Figure \ref{client}), \textit{DataWriter} (implementation of  OutputStream) contains a \textit{ByteBuffer}. When a higher layer method writes data to \textit{DataOuput}, \textit{DataWriter} put those bytes to its \textit{ByteBuffer}. When a selection event occurs, \textit{Datawriter} writes its \textit{ByteBuffer} to underline \textit{SocketChannel}. At server side (Figure \ref{server}), \textit{DataReader} contains a \textit{ByteBuffer} similar to client side. When a selection event occurs \textit{DataReader} input bytes from the \textit{SocketChannel} to its \textit{ByteBuffer}. When a \textit{ServerTask} reads data through \textit{DataInput}, it reads data from the \textit{ByteBuffer} and passes to higher layer.

\subsection{Load balancing}
Load balancing is the primary means of achieving data parallelism in stream processing. Each stage of the processing graph can be deployed into several nodes and can balance the load to next stage. In our framework load balancing happens at the \textit{Stream} level specific to the stream. \textit{KeyStream} balances the load by distributing keys among the nodes. \textit{RandomStream} balance messages randomly among next level of nodes. As mentioned above, buffering of the messages happens at the \textit{Stream} level. Further users can plug any custom stream with any specific partitioning function as well.

\subsection{Fault tolerance}
Fault tolerance is the ability of a system to operate with the presence of failures. These failures can be node failures or network failures and requirements to operate are depend on the system. In our framework, we provide fault resilient or ability to balance the load for existing nodes in the presence of node failures. If a receiving node fails then there will be a failure at the TCP connection and we close such connections and remove nodes from the \textit{Streams} (Stream objects hold the receiving node details). This prevents further routing messages to failure nodes.

### 2015-03-04 paper/system_overview.tex
updating the paper

\section{System Overview}
Our system mainly consists of a set of processes where each process receives events, processes them and forwards them to the next stage. We use the term adapters for special processes that receive events from external systems. These processes and adapters form a graph, which represents how a particular event stream is processed within the system. Our current implementation does not support fault tolerance and has adapted topology concept from Twitter Storm \cite{twitterStorm} and cluster concept from Yahoo S4\cite{neumeyer2010s4}. Rest of this section further describes this concept at a user level. In the next section we provide a detailed explanation about how we have made this inter-process communication efficient. 
\subsection{Process Graph}
A Process graph (Figure \ref{processgraph}) mainly consists of adapters, processes that execute the processing logic, and their interaction patterns. Adapters receive events from external systems and forward them to other processes. Processes process events according to a given logic and emit the generated events either to other processes or to outside systems.

\begin{figure}[!t]
	\centering
	\includegraphics[width=3.0in]{processgraph.png}
	\caption{Process Graph}
	\label{processgraph}
\end{figure}

\subsection{Runtime Graph}
When the system is deployed, there can be multiple instances of a particular stage to handle higher loads making the system scalable. Figure \ref{runtimegraph} shows possible runtime graph for above process graph. Here each parent process instance sends events to two child process instances. However in this case parent process instance has to pick a child process instance to send the message as well. One way of handling this problem is to pick a random node. Another approach is to send messages with same key to same child process instance.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{runtimegraph.png}
        \caption{Runtime Graph}
        \label{runtimegraph}
\end{figure}

@@

 The \texttt{Event} interface has a method to get the key for a given event. This key is used in sending messages to next processes as described earlier. \texttt{Parse} and \texttt{serialize} methods of \texttt{Event} interface are used to directly serialize the event attributes to underlying streams. When implementing these methods developers can define attribute serialization order in their \texttt{serialize} method and use the same order during \texttt{parse} method to avoid metadata passing with the binary format. \texttt{Processor} interface contains a method called \texttt{onEvent} which is invoked by the underlying framework when it receives an event to that process. The \texttt{start} method of the \texttt{Adapter} interface is invoked by the framework and that can be used to pull events from external systems once the system starts. During initialization both processes and adapters receive the container that can be used to send events to other processes.

@@

As shown in Figure \ref{deployment}, a deployment of the system consists of a manager node and a set of worker nodes. Worker nodes are grouped into clusters so that the application users can specify to which clusters each processes needs to get deployed. The Manager node processes applications, deploys them to worker nodes, and initiates the processing by starting the adapters.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{deployment.png}
        \caption{Deployment}
        \label{deployment}
\end{figure}

@@

\lstinputlisting[label=sampleruntime,caption=Runtime Graph,language=java]{runtimegraph.json}

>>>>

\section{System Overview}
Our system mainly consists of a set of processes where each process receives events, processes them and forwards them to the next stage. We use the term adapters for special processes that receive events from external systems. These processes and adapters form a graph, which represents how a particular event stream is processed within the system. Our current implementation has adapted the topology concept from Twitter Storm \cite{toshniwal2014storm} and the cluster concept from Yahoo S4 \cite{neumeyer2010s4}. The remaining of this section describes this concept at a user level. In the next section we provide a detailed explanation about how we have made the inter-process communication efficient. 
\subsection{Process Graph}
A Process graph mainly consists of adapters, processes that execute the processing logic, and their interaction patterns. Adapters receive events from external systems and forward them to other processes. Processes process events according to a given logic and emit the generated events either to other processes or to outside systems. Figure \ref{processgraph} shows an example process graph with 3 nodes. \textit{EventProducer} reads data from an external system and emits row events to \textit{EventParser}. \textit{Eventparser} processes the event and sends each word as an event. \textit{WordCounter} receives each word and emits the each word count periodically.

\begin{figure}[!t]
	\centering
	\includegraphics[width=3.0in]{processgraph.png}
	\caption{An example process graph with 3 nodes to count words.}
	\label{processgraph}
\end{figure}

\subsection{Runtime Graph}
When the system is deployed, there can be multiple instances of a particular stage to handle higher loads making the system scalable. Figure \ref{runtimegraph} shows possible runtime graph for process graph shown in Figure \ref{processgraph}. Here each parent process instance sends events to two child process instances. However in this case parent process instance has to pick a child process instance to send the message as well. One way of handling this problem is to pick a random node. Another approach is to send messages with same key to same child process instance.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{runtimegraph.png}
        \caption{Runtime graph for process graph given in Figure \ref{processgraph}.}
        \label{runtimegraph}
\end{figure}

@@

 The \texttt{Event} interface has a method to get the key for a given event. This key is used in sending messages to next processes as described earlier. \texttt{Serialize} method is used to directly convert the event attributes to binary format and \texttt{parse} method is used to directly retrieve event attribute values from binary format. When implementing these methods developers can define attribute serialization order in their \texttt{serialize} method and use the same order during \texttt{parse} method to avoid metadata passing with the binary format. \texttt{Processor} interface contains a method called \texttt{onEvent} which is invoked by the underlying framework when it receives an event to that process. The \texttt{start} method of the \texttt{Adapter} interface is invoked by the framework and that can be used to pull events from external systems once the system starts. During initialization both processes and adapters receive the container that can be used to send events to other processes.

@@

As shown in Figure \ref{deployment}, a deployment of the system consists of a manager node and a set of worker nodes. Worker nodes are grouped into clusters so that the application users can specify to which clusters each process needs to get deployed. The Manager node processes applications, deploys them to worker nodes, and initiates the processing by starting the adapters.

@@

\clearpage

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{deployment.png}
        \caption{Deployment view of the system with 6 worker nodes into 3 clusters and a Manager node.}
        \label{deployment}
\end{figure}

\lstinputlisting[captionpos=b, label=sampleruntime,caption=Configuration for runtime graph given in Figure \ref{runtimegraph},language=java]{runtimegraph.json}
